---
title: "Colour Modelling"
date: "`r Sys.Date()`"
author: "James Foster"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
    fig_caption: true
editor_options: 
  markdown: 
    wrap: sentence
---

## Details

> **DESCRIPTION**
>
> Models photoreceptor responses and spectral sensitivities using theoretical visual pigment templates.
> Simulates UV and green photoreceptor activation curves under experimental LED illumination to estimate relative photoreceptor excitation and the effects of unequal absolute sensitivity.

> **INPUTS**
>
> None (self-contained theoretical model)

> **OUTPUTS**
>
> `GUV_colour_modelling.html` (rendered report with figures)

> **REFERENCES**
>
> Menzel, R., Ventura, D. F., Hertel, H., de Souza, J. M. and Greggers, U. (1986).
> Spectral sensitivity of photoreceptors in insect eyes.
> *J. Comp. Physiol. A* 158, 361–377.
>
> Stavenga, D. G. (2010).
> On visual pigment templates and the spectral shape of invertebrate rhodopsins and metarhodopsins.
> *J. Comp. Physiol. A* 196, 869–878.
>
> Kolyfetis, G. *et al.* (2025).
> Electrophysiological recordings reveal photoreceptor coupling in the dorsal rim areas of honeybee and bumblebee eyes.
> *Biol. Lett.* 21:20250234.https://doi.org/10.1098/rsbl.2025.0234
>
> Backhaus W. (1992)
> The Bezold-Brücke effect in the color vision system of the honeybee.
> *Vision Res.* 32(8):1425-31. doi: 10.1016/0042-6989(92)90199-s

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document models photoreceptor responses and sensitivities based on the visual pigment templates described by Stavenga (2010) and physiological measurements of honeybee photoreceptors.

```{r setup_packages}
# Package for integration of functions
require("sfsmisc")
# Package for colour modelling
require("pavo")
```


## Retina-level model


### Photoreceptor Sensitivity

We define a function, `StavengaSpline()`, to generate visual pigment template curves for invertebrates.
It supports $A1$ pigment types and calculates the $\alpha$-, $\beta$-, and $\gamma$-absorption bands using modified lognormal models.

```{r stavenga_spline}
# Make a spline template for a visual pigment
StavengaSpline = function(spec_range = c(300, 700), #bounds of spectrum in nanometers
                          lambda_max,#peak sensitivity
                          a_type = 'a1'){#pigment type, only a1 available currently
  wlns =  seq(from = min(spec_range),
              to = max(spec_range), 
              length.out = 1e3) #
  # modified lognormal
  Mlognorm  = function(wl,l_max, a0, a1)
  {
    x = log10(wl/l_max)
    return(
      exp(-a0*x^2 *
            (1+a1*x+3*a1^2*x^2)
      )
    )
  }
  if(a_type == 'a1')
  {
    #alpha band
    a_band = Mlognorm(wl = wlns,
                      l_max = lambda_max, 
                      a0 = 380, 
                      a1 = 6.09)
    #beta band
    b_band = 0.29*Mlognorm(wl = wlns, 
                           l_max = 340, 
                           a0 = 247, 
                           a1 = 3.59)
    #gamma band
    g_band = 1.99*Mlognorm(wl = wlns, 
                           l_max = 276, 
                           a0 = 647, 
                           a1 = 23.4)
  }else
  {stop('a2 and a3 pigments not yet implemented')}
  # N.B. Stavenga normalises to max(a.band), I normalise to function max
  r_stav  = (a_band + b_band + g_band)/
    max(a_band + b_band + g_band)
  return(    smooth.spline(x = wlns,
                           y =  r_stav)    )
}#StavengaSpline <- function(spec.range, lambda.max)
```


### Response Function

We model the photoreceptor response (in mV) to light intensity (photons/cm²/s) using a Naka-Rushton-like function.
The parameters are based on honeybee photoreceptor recordings from Kolyfetis *et al.* (2025).
See `1Data/250327_009_Apis_MainRetina_Kolyfetis2025.pdf`.

```{r photoreceptor_response}
# Define a function for photoreceptor response
#params based on 250327_006 from Kolyfetis et al., 2025
PRresponse = function(
            quanta = 10^seq(from = 10, 
                            to  = 14, 
                            length.out = 21), #photons/cm2/s
            vmax = 50,#mV
            k50 = 0.03*1e14,#inflection point
            n = 1.07)#slope
  {
mV = (vmax*quanta^n)/ 
      (k50^n + quanta^n)
}
#this sets sensitivity in the following way
## rel_quanta = quanta/max(quanta)
## yy = mV/vmax
## sensitivity = k50*(yy/(1-yy))^(n^-1)
```


We establish parameters for the response peaks of honeybee photoreceptors (Menzel et al., 1986) and the emission peaks for the experimental ultraviolet and green LEDs, using a 20nm bandwidth range for the LED spectral distribution.

```{r wavelengths_of_interest}
# Response peak for honeybee photoreceptors Menzel et al., 1986
lmax_u = 343
lmax_g = 543
# Emission peak for each LED
led_u = 365
led_g = 528
# A 20nm (uniform) band for each 
wlband = seq(from = -10, 
             to  = 10, 
             by = 0.1)
```


### Sensitivity Curve Predictions

We predict the photoreceptors' sensitivity curves across the insect-visible spectrum and specifically for the LED emission bands.
We also calculate the integrated sensitivity for the green photoreceptor under green and UV LEDs to find its relative UV sensitivity (`g_rel_u`).

```{r predict_sensitivities}
# Predict sensitivity curves across whole spectrum
u_photopig = predict(StavengaSpline(lambda_max = lmax_u))
g_photopig = predict(StavengaSpline(lambda_max = lmax_g))

# Specifically for the LEDs
#green photoreceptor
gPR_uLED = predict(StavengaSpline(lambda_max = lmax_g), x = led_u + wlband)
gPR_gLED = predict(StavengaSpline(lambda_max = lmax_g), x = led_g + wlband)
#UV photoreceptor
uPR_gLED = predict(StavengaSpline(lambda_max = lmax_u), x = led_g + wlband)
uPR_uLED = predict(StavengaSpline(lambda_max = lmax_u), x = led_u + wlband)

# integrated sensitivity
sens_gPR_uLED = integrate.xy(gPR_uLED)
sens_gPR_gLED = integrate.xy(gPR_gLED)
g_rel_u = sens_gPR_uLED / sens_gPR_gLED
```

### Set Up Response Curve Data

We define a range of UV intensities (photons/cm²/s) and compute simulated response curves for the UV and green photoreceptors under different absolute sensitivity assumptions.

```{r setup_response_data}
xx = 10^seq(from = 11, to  = 15, length.out = 1e3) #photons/cm2/s
example_g_equal = PRresponse(k50 = 3e12/g_rel_u, quanta = xx)
example_g_lows = PRresponse(k50 = 3e13/g_rel_u, quanta = xx)
example_u = PRresponse(k50 = 3e12, quanta = xx)

#set margins all around for individual plots
par(mar = c(4.5,4.5,4.5,4.5))
```

### Equal Absolute Sensitivity

First, we plot simulated response curves assuming the green and UV photoreceptors have equal absolute sensitivity.
We also compute and overlay the log ratio of green to UV photoreceptor responses.

```{r plot_equal_sensitivity}
#open plot with green photoreceptor
plot(x = xx,
     y = example_g_equal,
     type = 'l',
     col = 'green',
     lwd = 5,
     log = 'x',
     ylim = c(0, 50),
     main = 'equal absolute sensitivity',
     xlab = 'UV intensity (photons/cm2/s)',
     ylab = 'photoreceptor response (mV)')
#add UV photoreceptor
lines(x = xx,
     y = example_u,
     col = 'magenta',
     lwd = 5)
#add reference lines
abline(h = c(0,0.5,1.0)*50,
       v = 10^c(11:15),
       col = 'gray')
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'))
#label curves
text(x = 10^c(12, 13.5, 13.7),
     y = c(25, 25, 30),
     labels = c('UV\nphotor.',
                'green\nphotor.',
                'log ratio'),
     col = c('magenta', 'green', 'purple4')
     )
#add log ratio
#open new plot
par(new = TRUE)
plot(x = xx,
     y = log(example_g_equal/example_u),
     type = 'l',
     ylim = c(-1.55, 0),
     log = 'x',
     lwd = 5,
     col = 'purple4',
     axes = FALSE,
     xlab = '',
     ylab = '')
#add axis
axis(side = 4,
     col = 'purple4')
#add axis label
mtext(side = 4,
      line = 2,
      text = 'log(green/UV)',
      col = 'purple4')
```

### Account for Sensitivity Difference

Next, we plot the response curves under a scenario where the green photoreceptor has a lower absolute sensitivity (semi-saturation constant $k_{50}$ at a higher intensity) than the UV photoreceptor.

```{r plot_lower_sensitivity}
plot(x = xx,
     y = example_g_lows,
     type = 'l',
     col = 'green',
     lwd = 5,
     log = 'x',
     ylim = c(0, 50),
     main = 'green lower absolute sensitivity',
     xlab = 'UV intensity (photons/cm2/s)',
     ylab = 'photoreceptor response (mV)')
lines(x = xx,
     y = example_u,
     col = 'magenta',
     lwd = 5)
#add reference lines
abline(h = c(0,0.5,1.0)*50,
       v = 10^c(11:15),
       col = 'gray')
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'))
#label curves
text(x = 10^c(12.2, 14.4, 13.7),
     y = c(25, 25, 25),
     labels = c('UV\nphotor.',
                'green\nphotor.',
                'log ratio'),
     col = c('magenta', 'green', 'purple4')
)
#add log ratio
#open new plot
par(new = TRUE)
plot(x = xx,
     y = log(example_g_lows/example_u),
     type = 'l',
     ylim = c(-4.0, 0),
     log = 'x',
     lwd = 5,
     col = 'purple4',
     axes = FALSE,
     xlab = '',
     ylab = '')

axis(side = 4,
     col = 'purple4')
mtext(side = 4,
      line = 2,
      text = 'log(green/UV)',
      col = 'purple4')
```

This relationship could also explain the change in perceived greenness of the bright UV stimulus, 
but would not explain the difference between dim and dimmer UV. 
Green receptors are also more common in the honeybee retina than UV receptors ($\approx 4.4:1$), 
which may somewhat compensate for any difference in sensitivity.


### Photoreceptor Contrast
Plot the simpler example to show how the green-opsin's $\beta$ peak and saturation of the UV photoreceptor could 
contribute to a change in perceived green-ness between our dim and bright UV stimuli.

This combined two-panel figure presents: 

1. 
The relative photoreceptor sensitivity profiles for both UV and green photoreceptors with integrated shaded areas for the UV and green LEDs.


2. 
The simulated response functions for both UV and green photoreceptors under the equal absolute sensitivity assumption, alongside the ratio of green to UV response on the right-hand axis.

```{r figure_4, fig.width=10, fig.height=5}
par(mfrow = c(1,2),
    mar = c(5,5,0,3))
plot(u_photopig,
     type = 'l',
     lwd = 5,
     col = 'magenta',
     xlim = c(300, 700),
     ylim = c(0,1),
     xlab = 'wavelength (nm)',
     ylab = 'photoreceptor sensitivity\n(relative)',
     )
lines(g_photopig,
      lwd = 5,
      col = 'green')

with(uPR_uLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('magenta',
                            alpha.f = 0.2),
          border = NA)
)
with(uPR_gLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('magenta',
                            alpha.f = 0.2),
          border = NA)
)

with(gPR_uLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('green',
                            alpha.f = 0.2),
          border = NA)
)
with(gPR_gLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('green',
                            alpha.f = 0.2),
          border = NA)
)
abline(h = c(0,1.0),
       v = 100*c(3:7),
       col = 'gray')


plot(x = xx,
     y = example_g_equal,
     type = 'l',
     col = 'green',
     lwd = 5,
     log = 'x',
     ylim = c(0, 50),
     xlab = 'UV intensity (photons/cm2/s)',
     ylab = 'photoreceptor response (mV)')
lines(x = xx,
      y = example_u,
      col = 'magenta',
      lwd = 5)
#add reference lines
abline(h = c(0,0.5,1.0)*50,
       v = 10^c(11:15),
       col = 'gray')
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'))
#label curves
text(x = 10^c(12.2, 14.4, 13.7),
     y = c(25, 25, 25),
     labels = c('UV\nphotor.',
                'green\nphotor.',
                'ratio'),
     col = c('magenta', 'green', 'purple4')
)
#add ratio
#open new plot
par(new = TRUE)
plot(x = xx,
     y = example_g_equal/example_u,
     type = 'l',
     ylim = c(0, 1),
     log = 'x',
     lwd = 5,
     col = 'purple4',
     axes = FALSE,
     xlab = '',
     ylab = '')

axis(side = 4,
     col = 'purple4')
mtext(side = 4,
      line = 2,
      text = 'green/UV',
      col = 'purple4')
```


## Colour model

### Set up stimuli

Here we construct our LED spectra as `rspec`` format spectra in µmol/m²/s, 
by converting from cm$^2$ to m$^2$ and photons to moles, multiplying by $10^6$ to get µmol.
For colour modelling, we need average photons per ommatidium, which requires multiplication 
by an additional scaling factor.

```{r setup-stimulus-spectrum}
LED_rspec = data.frame(wl = seq(from = 300, to = 700, by = 1.0))

stimulus_deg = 10.0   # Diameter of LED stimulus in degrees
acceptance_deg  = 2.0  # Acceptance angle of a honeybee ommatidium (~1.5° - 2.5°)

spatial_factor  =  (min(acceptance_deg, stimulus_deg) / stimulus_deg)^2 
scale_factor  = spatial_factor

micromol_stim = c(dimmer = 1.2e12, dim = 4e12, bright = 4e14) *
                 1e4 *       # cm2 -> m2
                 1e6 /       # mol -> micromol
                 6.022e23    # Avogadro's number

# Create uniform binary version
LED_rspec_rect = within(LED_rspec, {
  LEDu = ifelse(test = wl > min(led_u + wlband) & wl < max(led_u + wlband), yes = 1, no = 0)
  LEDg = ifelse(test = wl > min(led_g + wlband) & wl < max(led_g + wlband), yes = 1, no = 0)
})

# Apply spatial scaling and adjust by bandwidth
LED_rspec_rect = within(LED_rspec_rect, {
  bandwidth = diff(range(wlband))
  UVdimmer = (micromol_stim['dimmer'] * scale_factor) * LEDu / bandwidth
  UVdim    = (micromol_stim['dim']    * scale_factor) * LEDu / bandwidth
  UVbright = (micromol_stim['bright'] * scale_factor) * LEDu / bandwidth
  Gdim     = (micromol_stim['dim']    * scale_factor) * LEDg / bandwidth
  Gbright  = (micromol_stim['bright'] * scale_factor) * LEDg / bandwidth
  rm('LEDu', 'LEDg', 'bandwidth')
})

# Convert to standard pavo rspec format
LED_rspec_rect = as.rspec(LED_rspec_rect)

# plot params
vls = unlist(LED_rspec_rect[-1])
mn = min(vls[vls > 0])

plot(
  x = NULL, 
  xlim = c(300, 700), 
  ylim = c(mn, max(vls)),
  log = 'y',
  xlab = 'Wavelength (nm)',
  ylab = 'Intensity (µmol/m²/s)',
  main = 'Simplified Stimuli Profiles - Rectangular Spectrum'
)

with(LED_rspec_rect, {
  # Plot uniform blocks using step lines ("s")
  lines(x = wl, y = UVbright + mn, type = 's', col = 'purple4', lwd = 2.5)
  lines(x = wl, y = Gbright + mn, type = 's', col = 'forestgreen', lwd = 2.5)
  lines(x = wl, y = UVdim + mn, type = 's', col = 'purple4', lty = 3, lwd = 2)
  lines(x = wl, y = Gdim + mn, type = 's', col = 'forestgreen', lty = 3, lwd = 2)
})

```

Although measured spectra are not available for these LEDs, they could be simulated as Cauchy PDF.

```{r smooth-stimulus-spectrum}
scale_factor = spatial_factor

# Generate Cauchy profiles (typical LED emission)
LED_rspec_smooth = within(LED_rspec, {
  LEDu = dcauchy(x = wl, location = led_u, scale = diff(range(wlband))/2) # FWHM = 2γ
  LEDg = dcauchy(x = wl, location = led_g, scale = diff(range(wlband))/2) # FWHM = 2γ
})

# Apply spatial scaling directly
LED_rspec_smooth = within(LED_rspec_smooth, {
  UVdimmer = (micromol_stim['dimmer'] * scale_factor) * LEDu
  UVdim    = (micromol_stim['dim']    * scale_factor) * LEDu
  UVbright = (micromol_stim['bright'] * scale_factor) * LEDu
  Gdim     = (micromol_stim['dim']    * scale_factor) * LEDg
  Gbright  = (micromol_stim['bright'] * scale_factor) * LEDg
  rm('LEDu', 'LEDg')
})

# Convert to standard pavo rspec format
LED_rspec_smooth = as.rspec(LED_rspec_smooth)

# Plot smooth spectra
vls = unlist(LED_rspec_smooth[-1])
mn = min(vls[vls > 0])

plot(
  x = NULL, 
  xlim = c(300, 700), 
  ylim = c(mn, max(vls)),
  log = 'y',
  xlab = 'Wavelength (nm)',
  ylab = 'Intensity (µmol/m²/s)',
  main = 'Simplified Stimuli Profiles - Smooth Spectrum'
)

with(LED_rspec_smooth, {
  # Bright Cauchy Profiles
  polygon(x = c(wl, rev(wl)), y = c(UVbright, rep(mn, times = length(wl))),
          col = adjustcolor('magenta', alpha.f = 0.1), border = 'magenta')
  
  polygon(x = c(wl, rev(wl)), y = c(Gbright, rep(mn, times = length(wl))),
          col = adjustcolor('green3', alpha.f = 0.1), border = 'green3')

  # Dim Cauchy Profiles
  polygon(x = c(wl, rev(wl)), y = c(Gdim, rep(mn, times = length(wl))),
          col = adjustcolor('darkgreen', alpha.f = 0.25), border = 'darkgreen')

  polygon(x = c(wl, rev(wl)), y = c(UVdim, rep(mn, times = length(wl))),
          col = adjustcolor('purple', alpha.f = 0.25), border = 'purple')
})
```



### Simulate quantum catches
To estimate perceived colours, we need to first compute photoreceptor responses, 
as a function of quantum catch and background illumination, 
Photoreceptor quantum catches and relative responses are calculated by `pavo::vismodel`, 
then mapped into Colour-Opponent Coding colour space (`colspace`) for *Apis mellifera*.

The COC colour space requires that responses scale relative to background adaptation, 
achieved via von Kries scaling so that hyperbolic-transformed quantum catch ($(E_i = Q_i / (Q_i + 1))$) 
for each stimulus is divided by that for each background. 
*N.B.* This does not account for photoreceptor saturation.


```{r simulate-catch}

sky_spectrum = sensdata(illum = "bluesky")
sky_spectrum = within(sky_spectrum,
                     {
                      bluesky_mol = 1e14*bluesky/#multiply by flux of a dim blue sky
                        sum(bluesky)* #divide by integral
                 1e4 *       # cm2 -> m2
                 1e6 /       # mol -> micromol
                 6.022e23    # Avogadro's number
                     }
                     )

shade_spectrum = sensdata(illum = "forestshade")
shade_spectrum = within(shade_spectrum,
                     {
                      forestshade_mol = 3e12*forestshade* #multiply by flux of a dim forest
                 1e4 *       # cm2 -> m2
                 1e6 /       # mol -> micromol
                 6.022e23    # Avogadro's number
                     }
                     )

#check that the spectra formats match
nrow(LED_rspec_rect) == nrow(sky_spectrum)
nrow(LED_rspec_rect) == nrow(shade_spectrum)

qcatches_dark = vismodel(
  rspecdata = LED_rspec_rect, # Wavelengths + emission intensities
  visual = "apis",              # Honeybee trichromatic photoreceptors
  qcatch = "Ei",           #Colour-opponent coding model requires hyperbolic photoreceptor responses
  achromatic = "l",             # Use Green receptor for brightness contrast
  illum = "ideal",              # These are emission spectra not reflectances
  bkg = shade_spectrum[,'forestshade_mol'],           # The photoreceptors are adapted to shade
  vonkries = TRUE,             # Ideally we would turn off adaptation; no background light exists
  relative = FALSE               # Estimate absolute catches
)

qcatches_light = vismodel(
  rspecdata = LED_rspec_rect, # Wavelengths + emission intensities
  visual = "apis",              # Honeybee trichromatic photoreceptors
  qcatch = "Ei",           #Colour-opponent coding model requires hyperbolic photoreceptor responses
  achromatic = "l",             # Use Green receptor for brightness contrast
  illum = "ideal",              # These are emission spectra not reflectances
  bkg = sky_spectrum[,'bluesky_mol'],           # The photoreceptors are adapted to blue sky
  vonkries = TRUE,             # Necessary for COC model
  relative = FALSE               # Estimate absolute catches
)

# Project catches into the 2D space
bee_space_dark = colspace(vismodeldata = qcatches_dark,
                     space = "coc")#Backhaus Colour-opponent coding
bee_space_light = colspace(vismodeldata = qcatches_light,
                     space = 'coc')#Backhaus Colour-opponent coding
```


### Plot stimuli

Counterintuitively the COC model in `pavo` finds only weak saturation for the dim stimuli,
since the modelled quantum catchs are very low. 
While this does produce a shift in saturation with intensity, the shift is to more saturated,
the reverse of that shown by Backhaus (1992).
```{r colour_space_stimuli}


#plot colour space ----
par(mfrow = c(1,2),
    mar = c(0,1,0,1))
#original scaled
cocplot(
  bee_space_dark, 
  xlim = c(-1,1)*3,
  ylim = c(-1,1)*3,
  pch =  c(Gbright = 4,
         Gdim = 3,
         UVbright = 4,
         UVdim = 3,
         UVdimmer = 1),              
  labels = TRUE,
  col = c(Gbright = 'green3',
         Gdim = 'darkgreen',
         UVbright = 'magenta',
         UVdim = 'purple',
         UVdimmer = "magenta4"), 
  bg = gray(1.0, alpha = 0.3),
  lwd = 2,
  cex = 1.8              # Size of the points
)
mtext(text = 'dark adapted (shade)',
      at = c(2,2),
      line = -2)

#original scaled
cocplot(
  bee_space_light,
  xlim = c(-1,1)*3,
  ylim = c(-1,1)*3,
  pch = c(Gbright = 4,
         Gdim = 3,
         UVbright = 4,
         UVdim = 3,
         UVdimmer = 1),               
  labels = TRUE,
  col = c(Gbright = 'green3',
         Gdim = 'darkgreen',
         UVbright = 'magenta',
         UVdim = 'purple',
         UVdimmer = "magenta4"), 
  bg = gray(1.0, alpha = 0.3),
  lwd = 2,
  cex = 1.8              # Size of the points
)
mtext(text = 'light adapted (blue sky)',
      at = c(2,2),
      line = -2)

legend(
  "topleft", 
  legend = c("bright green",  
             "dim green",     
             "bright UV",
             "dim UV",
             "dimmer UV"), 
  col = c(Gbright = 'green3',
         Gdim = 'darkgreen',
         UVbright = 'magenta',
         UVdim = 'purple',
         UVdimmer = "magenta4"), 
  pt.bg = gray(1.0, alpha = 0.5),
  pch = c(Gbright = 4,
         Gdim = 3,
         UVbright = 4,
         UVdim = 3,
         UVdimmer = 1), 
  pt.cex = 1.0,
  cex = 1.2,
  bty = "n"               # Removes the legend border box
)

```



## Bezold-Brücke model


Set up a function to calculate colour shifts that would result from photoreceptor saturation.

```{r saturation-function}
simulate_bezold_bruecke <- function(flux_range, sens_matrix, e_param, v_max) {
  
  # 1. Pre-allocate vectors for performance
  n_steps       <- length(flux_range)
  ds_shifts     <- numeric(n_steps)
  ds_saturation <- numeric(n_steps)
  greenness     <- numeric(n_steps)
  bee_x_track   <- numeric(n_steps)
  bee_y_track   <- numeric(n_steps)
  
  # 2. Establish the baseline chromatic signal at your lowest intensity step
  phi_baseline   <- flux_range[1] * sens_matrix["UVdimmer", ]
  mv_baseline    <- PRresponse(phi_baseline, vmax = v_max)
  chrom_baseline <- mv_baseline - mean(mv_baseline)
  
  # 3. Cache repetitive noise denominator math outside the loop
  den_param <- e_param[1]^2 * e_param[2]^2 +
               e_param[2]^2 * e_param[3]^2 +
               e_param[3]^2 * e_param[1]^2
  
  # 4. Single execution loop for all sensory metrics
  for(i in 1:n_steps) {
    # Generate raw millivolt receptor potentials [V_s, V_m, V_l]
    phi_current <- flux_range[i] * sens_matrix["UVdimmer", ]
    mv_current  <- PRresponse(phi_current, vmax = v_max)
    
    # Isolate pure chromatic signal by stripping mean luminance
    mean_mv      <- mean(mv_current)
    chrom_signal <- mv_current - mean_mv
    
    # Chromatic shift (difference between mean-subtracted current vs baseline)
    delta_chrom  <- chrom_signal - chrom_baseline
    
    # Green receptor contribution ratio
    greenness[i] <- mv_current["l"] / sum(mv_current)
    
    # --- RNL Distance Formulas ---
    delta_num_param <- (delta_chrom[2] - delta_chrom[3])^2 * e_param[1]^2 +
                       (delta_chrom[3] - delta_chrom[1])^2 * e_param[2]^2 +
                       (delta_chrom[1] - delta_chrom[2])^2 * e_param[3]^2
    
    num_param <- (chrom_signal[2] - chrom_signal[3])^2 * e_param[1]^2 +
                 (chrom_signal[3] - chrom_signal[1])^2 * e_param[2]^2 +
                 (chrom_signal[1] - chrom_signal[2])^2 * e_param[3]^2
                 
    ds_shifts[i]     <- sqrt(delta_num_param / den_param)
    ds_saturation[i] <- sqrt(num_param / den_param)
    
    # --- Backhaus COC Cartesian Matrix Projection ---
    # Scaled safely by v_max to keep coordinates within standard -12 to 12 limits
    # index 1 = s (UV), index 2 = m (Blue), index 3 = l (Green)
    bee_x_track[i] <- 14 * (chrom_signal[3] - chrom_signal[2]) / v_max
    bee_y_track[i] <- 11 * (chrom_signal[1] - (chrom_signal[2] + chrom_signal[3]) / 2) / v_max
  }
  
  # 5. Compile outputs into a cohesive data frame
  results_df <- data.frame(
    flux          = flux_range,
    ds_shift      = ds_shifts,
    ds_saturation = ds_saturation,
    greenness     = greenness,
    coc_x         = bee_x_track,
    coc_y         = bee_y_track
  )
  
  return(results_df)
}

```

In order to 
• Purpose: simulate an intensity sweep (Bezold–Brücke) for a 365 nm LED: get receptor sensitivity matrix from vismodel, map photon flux → receptor mV (PRresponse), compute JND shift (manual RNL), and plot intensity-dependent hue shift.
• Key inputs: LED_rspec (UVdimmer), PRresponse function, flux_range, sens_matrix, RNL noise params (n_ratios, weber).
• Outputs: ds_shifts vector and log-x plot showing predicted chromatic shift vs intensity.
• Checks: PRresponse must be defined and return consistent units; sens_matrix must have correct row name ("UVdimmer"); watch numeric stability for very large flux values and ensure weber/e_param choice is appropriate.

```{r model-saturation}

v_max = 50#mV

# --- Use pavo to get the relative spectral overlap ratios ---
stimulated_spectra = LED_rspec_rect[c('wl','UVdimmer')]
stimulated_spectra = as.rspec(within(stimulated_spectra, {
  UVdimmer = UVdimmer / #convert back to photon flux
    (
      1e4 *       # cm2 -> m2
        1e6 /       # mol -> micromol
        6.022e23    # Avogadro's number)]
    )
  #normalise
  UVdimmer = UVdimmer/sum(UVdimmer)
}))
# Get clean, un-normalized sensitivity weights (relative = FALSE, scale = 1)
# This extracts the exact proportion of light each receptor type captures per wavelength
base_weights = vismodel(
  rspecdata = stimulated_spectra,
  visual = "apis",
  achromatic = "l",
  illum = "ideal",
  bkg = 0,
  vonkries = FALSE,
  relative = FALSE,
  scale = 1
)

# Extract just the UV (s), Blue (m), and Green (l) sensitivity matrix
sens_matrix = as.matrix(base_weights[, c("s", "m", "l")])

# Compute receptor noise
# Standard neural noise factor for a single honeybee receptor channel (Vorobyev et al., 2001)
w_noise = 0.05 

# Relative receptor counts across the retina based on overall macro ratios:
# UV (s) = ~11.4%, Blue (m) = ~5.4%, Green (l) = ~83.2%
# Scaled relative to 1 single UV receptor unit:
n_uv    = 1.0
n_blue  = 0.47
n_green = 7.30

# Calculate RNL integrated channel noise thresholds
e_uv    = w_noise / sqrt(n_uv)
e_blue  = w_noise / sqrt(n_blue)
e_green = w_noise / sqrt(n_green)

# Combined into the ordered e_param vector matching loop indices: [1]=s, [2]=m, [3]=l
e_param <- c(e_uv, e_blue, e_green)

# --- Simulate the Bezold–Brücke intensity sweep ---
# We will scale a single 365nm UV LED across your experimental range
flux_range = 10^seq(from = 10,
                    to = 15,
                    length.out = 1e3) # photons/cm2/s
    # ds_shifts = numeric(length(flux_range))
    # ds_saturation = numeric(length(flux_range))
    # greenness = numeric(length(flux_range))  # Relative L-channel contribution
    # 
    # # Calculate the millivolt baseline at your lowest intensity (10^10 photons)
    # # Multiplying the flux by the sensitivity matrix distributes the photons across the channels
    # # --- PART 4: Compute the JND color shift correctly ---
    # 
    # # 1. Baseline chromatic signal (isolated from overall luminance)
    # phi_baseline = flux_range[1] * sens_matrix["UVdimmer", ]# * flux_to_ommatidium
    # mv_baseline  = PRresponse(quanta = phi_baseline,
    #                           vmax = v_max)
    # chrom_baseline = (mv_baseline - mean(mv_baseline)) # Isolate pure baseline hue

    # for(i in 1:length(flux_range)) {
    #   phi_current = flux_range[i] * sens_matrix["UVdimmer", ] #* flux_to_ommatidium
    #   mv_current  = PRresponse(phi_current) # [V_s, V_m, V_l]
    #   
    #   # 2. Isolate pure current hue by removing mean luminance
    #   mean_mv = mean(mv_current)
    #   chrom_signal = (mv_current - mean_mv)
    #   
    #   # 3. Chromatic shift is the difference between the 
    #   # MEAN-SUBTRACTED signals (Current Hue vs Baseline Hue)
    #   delta_chrom = chrom_signal - chrom_baseline
    #   
    #   # Ratio of L-voltage relative to overall voltage response across all channels
    #   greenness[i]  =  mv_current["l"] / sum(mv_current)
    #   
    #   # 4. Manual RNL distance formula for pure HUE SHIFT (Bezold-Brücke)
    #   delta_num_param <- (delta_chrom[2] - delta_chrom[3])^2 * e_param[1]^2 +
    #                      (delta_chrom[3] - delta_chrom[1])^2 * e_param[2]^2 +
    #                      (delta_chrom[1] - delta_chrom[2])^2 * e_param[3]^2
    #   
    #   # RNL Chromatic Saturation (Distance from origin/achromatic point)
    #   num_param <- (chrom_signal[2] - chrom_signal[3])^2 * e_param[1]^2 +
    #                (chrom_signal[3] - chrom_signal[1])^2 * e_param[2]^2 +
    #                (chrom_signal[1] - chrom_signal[2])^2 * e_param[3]^2
    #                
    #   den_param <- e_param[1]^2 * e_param[2]^2 +
    #                e_param[2]^2 * e_param[3]^2 +
    #                e_param[3]^2 * e_param[1]^2
    #                
    #   ds_shifts[i]     <- sqrt(delta_num_param / den_param)
    #   ds_saturation[i] <- sqrt(num_param / den_param)
    # }

sat_data <- simulate_bezold_bruecke(
  flux_range  = flux_range,
  sens_matrix = sens_matrix,
  e_param     = e_param,
  v_max       = v_max
)

par(mfrow = c(1,2))
# --- Plot Intensity-dependent Hue Shift ---
plot(
  x = NULL,
  log = 'x',
  type = "l", 
  xlab = "UV intensity (photons/cm2/s)",
  ylab = "Chromatic Distance Shift (JND)",
  main = "Predicted Bezold–Brücke Shift for 365nm LED",
  ylim = c(0,1)*v_max,
  xlim = range(flux_range)
)
abline(h = 0)
polygon(y = c(0,1,1,0),
        x = rev(sort(rep(range(flux_range)*c(0.1, 10),times = 2))),
        col = gray(0.5, 0.2),
        border = NA
        )
with(sat_data,
     {
lines(x = flux, 
  y = ds_shift,
    pch = 19,
  col = "purple4",
  lwd = 5
  )
     }
)
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'),
       lwd = 2)
# --- Plot Intensity-dependent Saturation Shift ---
plot(
  x = NULL,
  log = 'x',
  type = "l", 
  xlab = "UV intensity (photons/cm2/s)",
  ylab = "Distance to Achromatic Point (JND)",
  main = "Predicted Bezold–Brücke Shift for 365nm LED",
  ylim = c(0,1)*v_max,
  xlim = range(flux_range)
)
abline(h = 0)
polygon(y = c(0,1,1,0),
        x = rev(sort(rep(range(flux_range)*c(0.1, 10),times = 2))),
        col = gray(0.5, 0.2),
        border = NA
        )
with(sat_data,
     {
lines(x = flux, 
  y = ds_saturation,
    pch = 19,
  col = "purple4",
  lwd = 5
  )
lines(x = flux, 
  y = greenness*v_max,
    pch = 19,
  col = "green",
  lwd = 5
  )
     }
)
axis(side = 4,
     at = c(0,0.5,1.0)*v_max,
     labels = c(0,0.5,1.0))
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'),
       lwd = 2)

```
### Hue analysis

Plot the outcome in COC space
```{r coc-saturation-plot}
# Add coordinate arrays to tracking variables
bee_x_track <- numeric(length(flux_range))
bee_y_track <- numeric(length(flux_range))
# --- Insert this inside your existing for-loop ---
# Backhaus matrix transformation constants mapping to x and y coordinates
# chrom_signal[1] = s (UV), chrom_signal[2] = m (Blue), chrom_signal[3] = l (Green)
for(i in 1:length(flux_range)) {
  phi_current = flux_range[i] * sens_matrix["UVdimmer", ] #* flux_to_ommatidium
  mv_current  = PRresponse(phi_current) # [V_s, V_m, V_l]  
  
  mean_mv = mean(mv_current)
  chrom_signal = mv_current - mean_mv
  
bee_x_track[i] = 7 * (chrom_signal[3] - chrom_signal[2])/v_max
bee_y_track[i] = 11 * (chrom_signal[1] - (chrom_signal[2] + chrom_signal[3]) / 2)/v_max
}
# 1. Package the track into a compliant pavo colspace data frame
coc_trajectory <- data.frame(
  s = mv_current[1]/v_max, # Dummy placeholders to pass class verification
  m = mv_current[2]/v_max,
  l = mv_current[3]/v_max,
  x = bee_x_track,   # Your intensity-driven x coordinates
  y = bee_y_track    # Your intensity-driven y coordinates
)
class(coc_trajectory) <- c("colspace", "coc", "data.frame")

# 2. Open a fresh plotting canvas
# pavo::cocplot builds the traditional bee hexagon boundary template
pavo::cocplot(coc_trajectory, 
              main = "Intensity-dependent Bezold-Brücke Trajectory",
              col = "transparent") # Keeps points hidden initially to avoid clutter

# 3. Draw the trajectory as a continuous curving vector path
lines(x = bee_x_track, 
      y = bee_y_track, 
      col = "purple4", 
      lwd = 4)

# 4. Highlight specific key experimental intensity milestones on top of the line
# Step 1 (10^10 photons)
points(x = bee_x_track[1], y = bee_y_track[1], col = "black", pch = 19, cex = 2)
text(x = bee_x_track[1], y = bee_y_track[1], labels = " Baseline (Dim)", pos = 3, cex = 0.8)

# Final Step (10^15 photons)
end_idx <- length(flux_range)
points(x = bee_x_track[end_idx], y = bee_y_track[end_idx], col = "red", pch = 19, cex = 2)
text(x = bee_x_track[end_idx], y = bee_y_track[end_idx], labels = " Saturated (Bright)", pos = 1, cex = 0.8)

```