---
title: "Colour Modelling"
date: "`r Sys.Date()`"
author: "James Foster"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
    fig_caption: true
editor_options: 
  markdown: 
    wrap: sentence
---

## Details

> **DESCRIPTION**
>
> Models photoreceptor responses and spectral sensitivities using theoretical visual pigment templates.
> It simulates UV and green photoreceptor activation curves under experimental LED illumination to analyze relative photoreceptor excitation and the effects of unequal absolute sensitivity.

> **INPUTS**
>
> None (self-contained theoretical model)

> **OUTPUTS**
>
> `GUV_colour_modelling.html` (rendered report with figures)

> **REFERENCES**
>
> Menzel, R., Ventura, D. F., Hertel, H., de Souza, J. M.
> and Greggers, U.
> (1986).
> Spectral sensitivity of photoreceptors in insect eyes.
> *J. Comp. Physiol. A* 158, 361–377.
>
> Stavenga, D. G.
> (2010).
> On visual pigment templates and the spectral shape of invertebrate rhodopsins and metarhodopsins.
> *J. Comp. Physiol. A* 196, 869–878.
>
> Kolyfetis, G.
> *et al.* (2025).
> Electrophysiological recordings reveal photoreceptor coupling in the dorsal rim areas of honeybee and bumblebee eyes.
> Biol.
> Lett.
> 21:20250234.https://doi.org/10.1098/rsbl.2025.0234

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document models photoreceptor responses and sensitivities based on the visual pigment templates described by Stavenga (2010) and physiological measurements of honeybee photoreceptors.

```{r setup_packages}
# Package for integration of functions
require("sfsmisc")
# Package for colour modelling
require("pavo")
```


## Retina-level model

### Photoreceptor Sensitivity

We define a function, `StavengaSpline()`, to generate visual pigment template curves for invertebrates.
It supports $A1$ pigment types and calculates the $\alpha$-, $\beta$-, and $\gamma$-absorption bands using modified lognormal models.

```{r stavenga_spline}
# Make a spline template for a visual pigment
StavengaSpline = function(spec_range = c(300, 700), #bounds of spectrum in nanometers
                          lambda_max,#peak sensitivity
                          a_type = 'a1'){#pigment type, only a1 available currently
  wlns =  seq(from = min(spec_range),
              to = max(spec_range), 
              length.out = 1e3) #
  # modified lognormal
  Mlognorm  = function(wl,l_max, a0, a1)
  {
    x = log10(wl/l_max)
    return(
      exp(-a0*x^2 *
            (1+a1*x+3*a1^2*x^2)
      )
    )
  }
  if(a_type == 'a1')
  {
    #alpha band
    a_band = Mlognorm(wl = wlns,
                      l_max = lambda_max, 
                      a0 = 380, 
                      a1 = 6.09)
    #beta band
    b_band = 0.29*Mlognorm(wl = wlns, 
                           l_max = 340, 
                           a0 = 247, 
                           a1 = 3.59)
    #gamma band
    g_band = 1.99*Mlognorm(wl = wlns, 
                           l_max = 276, 
                           a0 = 647, 
                           a1 = 23.4)
  }else
  {stop('a2 and a3 pigments not yet implemented')}
  # N.B. Stavenga normalises to max(a.band), I normalise to function max
  r_stav  = (a_band + b_band + g_band)/
    max(a_band + b_band + g_band)
  return(    smooth.spline(x = wlns,
                           y =  r_stav)    )
}#StavengaSpline <- function(spec.range, lambda.max)
```


### Response Function

We model the photoreceptor response (in mV) to light intensity (photons/cm²/s) using a Naka-Rushton-like function.
The parameters are based on honeybee photoreceptor recordings from Kolyfetis *et al.* (2025).
See `1Data/250327_009_Apis_MainRetina_Kolyfetis2025.pdf`.

```{r photoreceptor_response}
# Define a function for photoreceptor response
#params based on 250327_006 from Kolyfetis et al., 2025
PRresponse = function(
            quanta = 10^seq(from = 10, 
                            to  = 14, 
                            length.out = 21), #photons/cm2/s
            vmax = 50,#mV
            k50 = 0.03*1e14,#inflection point
            n = 1.07)#slope
  {
mV = (vmax*quanta^n)/ 
      (k50^n + quanta^n)
}
#this sets sensitivity in the following way
## rel_quanta = quanta/max(quanta)
## yy = mV/vmax
## sensitivity = k50*(yy/(1-yy))^(n^-1)
```


We establish parameters for the response peaks of honeybee photoreceptors (Menzel et al., 1986) and the emission peaks for the experimental ultraviolet and green LEDs, using a 20nm bandwidth range for the LED spectral distribution.

```{r wavelengths_of_interest}
# Response peak for honeybee photoreceptors Menzel et al., 1986
lmax_u = 343
lmax_g = 543
# Emission peak for each LED
led_u = 365
led_g = 528
# A 20nm (uniform) band for each 
wlband = seq(from = -10, 
             to  = 10, 
             by = 0.1)
```


### Sensitivity Curve Predictions

We predict the photoreceptors' sensitivity curves across the insect-visible spectrum and specifically for the LED emission bands.
We also calculate the integrated sensitivity for the green photoreceptor under green and UV LEDs to find its relative UV sensitivity (`g_rel_u`).

```{r predict_sensitivities}
# Predict sensitivity curves across whole spectrum
u_photopig = predict(StavengaSpline(lambda_max = lmax_u))
g_photopig = predict(StavengaSpline(lambda_max = lmax_g))

# Specifically for the LEDs
#green photoreceptor
gPR_uLED = predict(StavengaSpline(lambda_max = lmax_g), x = led_u + wlband)
gPR_gLED = predict(StavengaSpline(lambda_max = lmax_g), x = led_g + wlband)
#UV photoreceptor
uPR_gLED = predict(StavengaSpline(lambda_max = lmax_u), x = led_g + wlband)
uPR_uLED = predict(StavengaSpline(lambda_max = lmax_u), x = led_u + wlband)

# integrated sensitivity
sens_gPR_uLED = integrate.xy(gPR_uLED)
sens_gPR_gLED = integrate.xy(gPR_gLED)
g_rel_u = sens_gPR_uLED/sens_gPR_gLED
```

### Set Up Response Curve Data

We define a range of UV intensities (photons/cm²/s) and compute simulated response curves for the UV and green photoreceptors under different absolute sensitivity assumptions.

```{r setup_response_data}
xx = 10^seq(from = 11, to  = 15, length.out = 1e3) #photons/cm2/s
example_g_equal = PRresponse(k50 = 3e12/g_rel_u, quanta = xx)
example_g_lows = PRresponse(k50 = 3e13/g_rel_u, quanta = xx)
example_u = PRresponse(k50 = 3e12, quanta = xx)

#set margins all around for individual plots
par(mar = c(4.5,4.5,4.5,4.5))
```

### Equal Absolute Sensitivity

First, we plot simulated response curves assuming the green and UV photoreceptors have equal absolute sensitivity.
We also compute and overlay the log ratio of green to UV photoreceptor responses.

```{r plot_equal_sensitivity}
#open plot with green photoreceptor
plot(x = xx,
     y = example_g_equal,
     type = 'l',
     col = 'green',
     lwd = 5,
     log = 'x',
     ylim = c(0, 50),
     main = 'equal absolute sensitivity',
     xlab = 'UV intensity (photons/cm2/s)',
     ylab = 'photoreceptor response (mV)')
#add UV photoreceptor
lines(x = xx,
     y = example_u,
     col = 'magenta',
     lwd = 5)
#add reference lines
abline(h = c(0,0.5,1.0)*50,
       v = 10^c(11:15),
       col = 'gray')
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'))
#label curves
text(x = 10^c(12, 13.5, 13.7),
     y = c(25, 25, 30),
     labels = c('UV\nphotor.',
                'green\nphotor.',
                'log ratio'),
     col = c('magenta', 'green', 'purple4')
     )
#add log ratio
#open new plot
par(new = TRUE)
plot(x = xx,
     y = log(example_g_equal/example_u),
     type = 'l',
     ylim = c(-1.55, 0),
     log = 'x',
     lwd = 5,
     col = 'purple4',
     axes = FALSE,
     xlab = '',
     ylab = '')
#add axis
axis(side = 4,
     col = 'purple4')
#add axis label
mtext(side = 4,
      line = 2,
      text = 'log(green/UV)',
      col = 'purple4')
```

### Account for Sensitivity Difference

Next, we plot the response curves under a scenario where the green photoreceptor has a lower absolute sensitivity (higher semi-saturation constant $k_{50}$) than the UV photoreceptor.

```{r plot_lower_sensitivity}
plot(x = xx,
     y = example_g_lows,
     type = 'l',
     col = 'green',
     lwd = 5,
     log = 'x',
     ylim = c(0, 50),
     main = 'green lower absolute sensitivity',
     xlab = 'UV intensity (photons/cm2/s)',
     ylab = 'photoreceptor response (mV)')
lines(x = xx,
     y = example_u,
     col = 'magenta',
     lwd = 5)
#add reference lines
abline(h = c(0,0.5,1.0)*50,
       v = 10^c(11:15),
       col = 'gray')
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'))
#label curves
text(x = 10^c(12.2, 14.4, 13.7),
     y = c(25, 25, 25),
     labels = c('UV\nphotor.',
                'green\nphotor.',
                'log ratio'),
     col = c('magenta', 'green', 'purple4')
)
#add log ratio
#open new plot
par(new = TRUE)
plot(x = xx,
     y = log(example_g_lows/example_u),
     type = 'l',
     ylim = c(-4.0, 0),
     log = 'x',
     lwd = 5,
     col = 'purple4',
     axes = FALSE,
     xlab = '',
     ylab = '')

axis(side = 4,
     col = 'purple4')
mtext(side = 4,
      line = 2,
      text = 'log(green/UV)',
      col = 'purple4')
```

### Photoreceptor Contrast

This combined two-panel figure presents: 

1. 
The relative photoreceptor sensitivity profiles for both UV and green photoreceptors with integrated shaded areas for the UV and green LEDs.


2. 
The simulated response functions for both UV and green photoreceptors under the equal absolute sensitivity assumption, alongside the ratio of green to UV response on the right-hand axis.

```{r figure_4, fig.width=10, fig.height=5}
par(mfrow = c(1,2),
    mar = c(5,5,0,3))
plot(u_photopig,
     type = 'l',
     lwd = 5,
     col = 'magenta',
     xlim = c(300, 700),
     ylim = c(0,1),
     xlab = 'wavelength (nm)',
     ylab = 'photoreceptor sensitivity\n(relative)',
     )
lines(g_photopig,
      lwd = 5,
      col = 'green')

with(uPR_uLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('magenta',
                            alpha.f = 0.2),
          border = NA)
)
with(uPR_gLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('magenta',
                            alpha.f = 0.2),
          border = NA)
)

with(gPR_uLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('green',
                            alpha.f = 0.2),
          border = NA)
)
with(gPR_gLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('green',
                            alpha.f = 0.2),
          border = NA)
)
abline(h = c(0,1.0),
       v = 100*c(3:7),
       col = 'gray')


plot(x = xx,
     y = example_g_equal,
     type = 'l',
     col = 'green',
     lwd = 5,
     log = 'x',
     ylim = c(0, 50),
     xlab = 'UV intensity (photons/cm2/s)',
     ylab = 'photoreceptor response (mV)')
lines(x = xx,
      y = example_u,
      col = 'magenta',
      lwd = 5)
#add reference lines
abline(h = c(0,0.5,1.0)*50,
       v = 10^c(11:15),
       col = 'gray')
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'))
#label curves
text(x = 10^c(12.2, 14.4, 13.7),
     y = c(25, 25, 25),
     labels = c('UV\nphotor.',
                'green\nphotor.',
                'ratio'),
     col = c('magenta', 'green', 'purple4')
)
#add ratio
#open new plot
par(new = TRUE)
plot(x = xx,
     y = example_g_equal/example_u,
     type = 'l',
     ylim = c(0, 1),
     log = 'x',
     lwd = 5,
     col = 'purple4',
     axes = FALSE,
     xlab = '',
     ylab = '')

axis(side = 4,
     col = 'purple4')
mtext(side = 4,
      line = 2,
      text = 'green/UV',
      col = 'purple4')
```


## Colour model

### Set up stimuli

```{r setup-stimulus-spectrum}
LED_rspec = data.frame(wl = seq(from = 300,
                                to = 700,
                                by = 1.0) # pavo only uses 1 nm increments
                       )
# LED_rspec = within(LED_rspec,
#                    {
#                      LEDu = ifelse(test = wl > min(led_u + wlband) & wl < max(led_u + wlband),
#                                    yes = 1,
#                                    no = 0
#                                    )
#                      LEDg = ifelse(test = wl > min(led_g + wlband) & wl < max(led_g + wlband),
#                                    yes = 1,
#                                    no = 0
#                                    )
#                    }
#                    )

#uniform spectra fail for colour calculations, use Cauchy (typical LED emission)
LED_rspec = within(LED_rspec,
                   {
                     LEDu = dcauchy(x = wl,
                                    location = led_u,
                                    scale = diff(range(wlband))/2 # for Cauchy FWHM = 2γ
                                   )
                     LEDg = dcauchy(x = wl,
                                    location = led_g,
                                    scale = diff(range(wlband))/2 # for Cauchy FWHM = 2γ
                                   )
                   }
                   )
#convert photon flux from photons/cm2/s to micro mol/m2/s
micromol_stim <- c(dimmer = 1.2e12, dim = 4e12, bright = 4e14) *
                    1e4 * #cm2 -> m2
                    1e6 / #mol -> micromol
                    6.022e23 # Avogadro's number

LED_rspec = within(LED_rspec,
                   {
                     UVdimmer = micromol_stim['dimmer']*
                       LEDu#/ # at those wavelengths in the LED emission halfmax
                       # diff(range(wlband)) #divide by width (nm)
                     UVdim = micromol_stim['dim']*LEDu#/
                       # diff(range(wlband))
                     UVbright = micromol_stim['bright']*LEDu#/
                       # diff(range(wlband))
                     Gdim = micromol_stim['dim']*LEDg#/
                       # diff(range(wlband))
                     Gbright = micromol_stim['bright']*LEDg#/
                       # diff(range(wlband))
                     rm('LEDu', 'LEDg')
                   }
                   )

with(LED_rspec,
     {
      plot(x = wl,
           y = UVbright,
           type = 'l',
           col = 'magenta',
           # log = 'y',
           xlab = 'wavelength (nm)',
           ylab = 'intensity (micro mol/m2/s)',
           main = 'Simplified stimuli')
       lines(x = wl,
           y = Gbright,
           col = 'green3')       
       # lines(x = wl,
       #     y = Gdim,
       #     col = 'darkgreen')
       # lines(x = wl,
       #     y = UVdim,
       #     col = 'purple')
       polygon(x = c(wl, rev(wl)),
           y = c(Gdim, rep(0, times = length(wl))),
           col = 'darkgreen',
           bg = adjustcolor('darkgreen', alpha.f = 0.7))
       polygon(x = c(wl, rev(wl)),
           y = c(UVdim, rep(0, times = length(wl))),
           col = 'purple',
           bg = adjustcolor('purple', alpha.f = 0.7))
     }
     )
```

### Simulate quantum catches

```{r simulate-catch}
qcatches = vismodel(
  rspecdata = LED_rspec, # Wavelengths + emission intensities
  visual = "apis",              # Honeybee trichromatic photoreceptors
  achromatic = "l",             # Use Green receptor for brightness contrast
  illum = "ideal",              # These are emission spectra not reflectances
  bkg = "ideal",                # Flat baseline (dark room)
  vonkries = FALSE,             # Turn off adaptation; no background light exists
  relative = FALSE               # Estimate absolute catches
)
# Project catches into the 2D space
bee_space = colspace(vismodeldata = qcatches)#use default colour space

# Calculate chromatic and achromatic distances between all stimuli
percept_distances = coldist(bee_space,
                            noise = "quantum")#minimal neural noise under dark conditions

# Inspect the distance matrix
print(percept_distances, digits = 3)

#increase quantum catch for JND comparison (too dark)
qcatches_brighter = vismodel(
  rspecdata = LED_rspec,# Wavelengths + emission intensities
  visual = "apis",              # Honeybee trichromatic photoreceptors
  achromatic = "l",             # Use Green receptor for brightness contrast
  illum = "ideal",              # These are emission spectra not reflectances
  bkg = 0,                # Surrounding environment absorbs all light
  vonkries = FALSE,             # Turn off adaptation; no background light exists
  relative = FALSE,# Estimate absolute catches
  scale = 1e6 # rescale to avoid log-error
)

bee_space_brighter = colspace(vismodeldata = qcatches_brighter)

# Calculate both chromatic and achromatic distances using receptor noise
bee_jnd = coldist(
  qcatches_brighter, # Pass the VISMODEL object, NOT the colspace object
  noise = "quantum",         # Dynamic photon-shot noise 
  achromatic = TRUE,         #  calculate dL
  n = c(1, 2, 6),            # Honeybee photoreceptor ratios (UV, Blue, Green)
  weber = 0.05, # Chromatic Weber (Osorio & Vorobyev)
  weber.achro = 0.1          # Achromatic Weber fraction for the green channel (Hempel de Ibarra et al., https://doi.org/10.1007/s003590100192)
)

# Print Just noticeable distances
print(bee_jnd, digits = 3)
```


### Plot stimuli

```{r colour_space_stimuli}


#plot colour space ----
par(mfrow = c(1,2),
    mar = c(0,1,0,1))
#original scaled
triplot(
  bee_space, 
  pch = 21,               # Filled circles with borders
  labels = TRUE,
  col = c(Gbright = 'green3',
         Gdim = 'darkgreen',
         UVbright = 'magenta',
         UVdim = 'purple',
         UVdimmer = "purple4"), 
  bg = gray(1.0, alpha = 0.3),
  lwd = 2,
  cex = 1.8,              # Size of the points
  gamut = TRUE # show possible colours
)


legend(
  "topleft", 
  legend = c("bright green",  
             "dim green",     
             "bright UV",
             "dim UV",
             "dimmer UV"), 
  col = c(Gbright = 'green3',
         Gdim = 'darkgreen',
         UVbright = 'magenta',
         UVdim = 'purple',
         UVdimmer = "purple4"), 
  pt.bg = gray(1.0, alpha = 0.5),
  pch = 21, 
  pt.cex = 1.0,
  cex = 1.2,
  bty = "n"               # Removes the legend border box
)


# plot perceived differences ----
par(mar = c(5,5,5,0))
with(bee_jnd,
     {
  plot(
    NULL,
    xlab = "Chromatic Distance (JND)",
    ylab = "Achromatic Contrast (JND)",
    main = "Perceptual Separation",
  xlim = c(0,max(dS)+10),
  ylim = c(0,max(dL)+5)
  )
# Add a threshold line at 1 JND (the theoretical limit of discrimination)
polygon(x = c(0,1,1,0),
        y = c(100,100,0,0),
        col = gray(0.5, 0.2),
        border = NA
        )
polygon(y = c(0,1,1,0),
        x = c(100,100,0,0),
        col = gray(0.5, 0.2),
        border = NA
        )       
       
points(
  x = dS, 
  y = dL,
  pch = 21,
  col = c(Gbright_Gdim =  'green',
  Gbright_UVbright = 'green',
  Gbright_UVdim =  'green',
  Gbright_UVdimmer = 'green',
  Gdim_UVbright = 'darkgreen',
  Gdim_UVdim = 'darkgreen',
  Gdim_UVdimmer = 'darkgreen',
  UVbright_UVdim = 'magenta',
  UVbright_UVdimmer = 'magenta',
  UVdim_UVdimmer = 'purple'),
  bg = c(Gbright_Gdim =  'darkgreen',
  Gbright_UVbright = 'magenta',
  Gbright_UVdim =  'purple',
  Gbright_UVdimmer = 'purple4',
  Gdim_UVbright = 'magenta',
  Gdim_UVdim = 'purple',
  Gdim_UVdimmer = 'purple4',
  UVbright_UVdim = 'purple',
  UVbright_UVdimmer = 'purple4',
  UVdim_UVdimmer = 'purple4'),
  cex = 1.5,
  lwd = 2
)



text(
  x = dS, 
  y = dL, 
  labels = paste(patch1, "vs", patch2), 
  pos = 4, 
  offset = 0.5, 
  cex = 0.75, 
  # font = 2,             # Bold text font
  col = "grey10"
)
     }
)

```


## Saturation model - WIP!

```{r model-saturation}

# --- PART 1: Use pavo to get the relative spectral overlap ratios ---
# Generate standard 1-nm curves
wl_vector <- 300:700
uv_source    <- dcauchy(wl_vector, location = 365, scale = 10)
green_source <- dcauchy(wl_vector, location = 528, scale = 10)

stimulated_spectra <- as.rspec(data.frame(wl = wl_vector, uv_led = uv_source, green_led = green_source))

# Get clean, un-normalized sensitivity weights (relative = FALSE, scale = 1)
# This extracts the exact proportion of light each receptor type captures per wavelength
base_weights <- vismodel(
  rspecdata = stimulated_spectra,
  visual = "apis", achromatic = "l", illum = "ideal", bkg = 0, vonkries = FALSE, relative = FALSE, scale = 1
)

# Extract just the UV (s), Blue (m), and Green (l) sensitivity matrix
sens_matrix <- as.matrix(base_weights[, c("s", "m", "l")])


# --- PART 2: Define your custom biological mV function ---
# Modify it slightly to accept your calculated physical photon flux scaling
PRresponse <- function(quanta, vmax = 50, k50 = 0.03 * 1e14, n_slope = 1.07) {
  mV <- (vmax * quanta^n_slope) / (k50^n_slope + quanta^n_slope)
  return(mV)
}


# --- PART 3: Simulate the Bezold–Brücke intensity sweep ---
# We will scale a single 365nm UV LED across your experimental range
flux_range <- 10^seq(from = 10, to = 14, length.out = 21) # photons/cm2/s
ds_shifts  <- numeric(length(flux_range))

# Calculate the millivolt baseline at your lowest intensity (10^10 photons)
# Multiplying the flux by the sensitivity matrix distributes the photons across the channels
phi_baseline <- flux_range[1] * sens_matrix["uv_led", ]
mv_baseline  <- PRresponse(phi_baseline)

# Standard apis RNL noise parameters
n_ratios <- c(1, 2, 6) # UV, Blue, Green receptor abundance
weber    <- 0.05
e        <- sqrt(weber^2 + 1/n_ratios)

# --- PART 4: Compute the JND color shift manually ---
for(i in 1:length(flux_range)) {
  # Calculate actual photons caught by each receptor type at this specific brightness
  phi_current <- flux_range[i] * sens_matrix["uv_led", ]
  
  # CRITICAL PHYSIOLOGY STEP: Convert the photons into true millivolts using your function
  mv_current  <- PRresponse(phi_current)
  
  # In visual ecology, when using real mV responses, the signal difference 
  # is modeled as the direct subtraction of millivolts divided by noise, 
  # bypassing pavo's log step because the Naka-Rushton curves ALREADY handle the scaling compression!
  delta_signal <- mv_current - mv_baseline
  
  # Manual RNL distance formula for trichromats
  num <- (delta_signal[2] - delta_signal[3])^2 * e[1]^2 + 
         (delta_signal[3] - delta_signal[1])^2 * e[2]^2 + 
         (delta_signal[1] - delta_signal[2])^2 * e[3]^2
  den <- e[1]^2 * e[2]^2 + e[2]^2 * e[3]^2 + e[3]^2 * e[1]^2
  
  ds_shifts[i] <- sqrt(num / den)
}

# --- PART 5: Plot your Intensity-dependent Hue Shift ---
plot(
  log10(flux_range), ds_shifts,
  type = "b", pch = 19, col = "purple",
  xlab = "Log10 Photon Flux (photons/cm²/s)",
  ylab = "Chromatic Distance Shift (dS in JNDs)",
  main = "Predicted Bezold–Brücke Shift for 365nm LED"
)
abline(h = 1, lty = 2, col = "red", lwd = 1.5) # The 1 JND threshold line

```