---
title: "Colour Modelling"
date: "`r Sys.Date()`"
author: "James Foster"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
    fig_caption: true
editor_options: 
  markdown: 
    wrap: sentence
---

## Details

> **DESCRIPTION**
>
> Models photoreceptor responses and spectral sensitivities using theoretical visual pigment templates. It simulates UV and green photoreceptor activation curves under experimental LED illumination to analyze relative photoreceptor excitation and the effects of unequal absolute sensitivity.

> **INPUTS**
>
> None (self-contained theoretical model)

> **OUTPUTS**
>
> `GUV_colour_modelling.html` (rendered report with figures)

> **REFERENCES**
>
> Kolyfetis, G. et al. (2025). Unpublished photoreceptor recordings (cell `250327_006`).
>
> Menzel, R., Ventura, D. F., Hertel, H., de Souza, J. M. and Greggers, U. (1986). Spectral sensitivity of photoreceptors in insect eyes. *J. Comp. Physiol. A* 158, 361–377.
>
> Stavenga, D. G. (2010). On visual pigment templates and the spectral shape of invertebrate rhodopsins and metarhodopsins. *J. Comp. Physiol. A* 196, 869–878.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document models photoreceptor responses and sensitivities based on the visual pigment templates described by Stavenga (2010) and physiological measurements of honeybee photoreceptors.

```{r setup_libs, include=FALSE}
# Load required library for integration functions
library(sfsmisc)
```

## Photoreceptor Sensitivity (Stavenga Spline)

We define a function, `StavengaSpline()`, to generate visual pigment template curves for invertebrates. It supports $A1$ pigment types and calculates the $\alpha$-, $\beta$-, and $\gamma$-absorption bands using modified lognormal models.

*Reference:* Stavenga, D. G. (2010). On visual pigment templates and the spectral shape of invertebrate rhodopsins and metarhodopsins. *Journal of Comparative Physiology A*, 196(11), 869–878.

```{r stavenga_spline}
# Make a spline template for a visual pigment
StavengaSpline = function(spec_range = c(300, 700), #bounds of spectrum in nanometers
                          lambda_max,#peak sensitivity
                          a_type = 'a1'){#pigment type, only a1 available currently
  wlns =  seq(from = min(spec_range),
              to = max(spec_range), 
              length.out = 1e3) #
  #Stavenga, D. G. (2010). On visual pigment templates and the spectral shape of invertebrate rhodopsins and metarhodopsins. Journal of Comparative Physiology A: Neuroethology, Sensory, Neural, and Behavioral Physiology, 196(11), 869–878. doi:10.1007/s00359-010-0568-7
  # modified lognormal
  Mlognorm  = function(wl,l_max, a0, a1)
  {
    x = log10(wl/l_max)
    return(
      exp(-a0*x^2 *
            (1+a1*x+3*a1^2*x^2)
      )
    )
  }
  if(a_type == 'a1')
  {
    #alpha band
    a_band = Mlognorm(wl = wlns,
                      l_max = lambda_max, 
                      a0 = 380, 
                      a1 = 6.09)
    #beta band
    b_band = 0.29*Mlognorm(wl = wlns, 
                           l_max = 340, 
                           a0 = 247, 
                           a1 = 3.59)
    #gamma band
    g_band = 1.99*Mlognorm(wl = wlns, 
                           l_max = 276, 
                           a0 = 647, 
                           a1 = 23.4)
  }else
  {stop('a2 and a3 pigments not yet implemented')}
  # N.B. Stavenga normalises to max(a.band), I normalise to function max
  r_stav  = (a_band + b_band + g_band)/
    max(a_band + b_band + g_band)
  return(    smooth.spline(x = wlns,
                           y =  r_stav)    )
}#StavengaSpline <- function(spec.range, lambda.max)
```

## Photoreceptor Response Function

We model the photoreceptor response (in mV) to light intensity (photons/cm²/s) using a Naka-Rushton-like function. The parameters are based on honeybee photoreceptor recordings from Kolyfetis et al. (2025).

*Reference Spreadsheet:* [Kolyfetis et al. (2025) Data](https://docs.google.com/spreadsheets/d/18r4MJe30J2Hgh7MQHX0us8KqTKhufxmgN_C4IpbRGmc/edit?gid=2023063010#gid=2023063010&range=A1:F1)

```{r photoreceptor_response}
# Define a function for photoreceptor response
#params based on 250327_006 from Kolyfetis et al., 2025
PRresponse = function(
            quanta = 10^seq(from = 10, to  = 14, length.out = 21), #photons/cm2/s
            vmax = 50,#mV
            k50 = 0.03*1e14,#inflection point
            n = 1.07)#slope
  {
mV = (vmax*quanta^n)/ 
      (k50^n + quanta^n)
}
#this sets sensitivity in the following way
## rel_quanta = quanta/max(quanta)
## yy = mV/vmax
## sensitivity = k50*(yy/(1-yy))^(n^-1)
```

## Wavelengths of Interest

We establish parameters for the response peaks of honeybee photoreceptors (Menzel et al., 1986) and the emission peaks for the experimental ultraviolet and green LEDs, using a 20nm bandwidth range for the LED spectral distribution.

```{r wavelengths_of_interest}
# Response peak for honeybee photoreceptors Menzel et al., 1986
lmax_u = 343
lmax_g = 543
# Emission peak for each LED
led_u = 365
led_g = 528
# A 20nm (uniform) band for each 
wlband = seq(from = -10, to  = 10, by = 0.1)
```

## Sensitivity Curve Predictions

Using our Stavenga Splines, we predict the photoreceptors' sensitivity curves across the entire spectrum and specifically for the LED emission bands. We also calculate the integrated sensitivity for the green photoreceptor under green and UV LEDs to find its relative UV sensitivity (`g_rel_u`).

```{r predict_sensitivities}
# Predict sensitivity curves across whole spectrum
u_photopig = predict(StavengaSpline(lambda_max = lmax_u))
g_photopig = predict(StavengaSpline(lambda_max = lmax_g))

# Specifically for the LEDs
#green photoreceptor
gPR_uLED = predict(StavengaSpline(lambda_max = lmax_g), x = led_u + wlband)
gPR_gLED = predict(StavengaSpline(lambda_max = lmax_g), x = led_g + wlband)
#UV photoreceptor
uPR_gLED = predict(StavengaSpline(lambda_max = lmax_u), x = led_g + wlband)
uPR_uLED = predict(StavengaSpline(lambda_max = lmax_u), x = led_u + wlband)

# integrated sensitivity
sens_gPR_uLED = integrate.xy(gPR_uLED)
sens_gPR_gLED = integrate.xy(gPR_gLED)
g_rel_u = sens_gPR_uLED/sens_gPR_gLED
```

## Set Up Response Curve Data

We define a wide range of UV intensities (photons/cm²/s) and compute simulated response curves for the UV and green photoreceptors under different absolute sensitivity assumptions.

```{r setup_response_data}
xx = 10^seq(from = 11, to  = 15, length.out = 1e3) #photons/cm2/s
example_g_equal = PRresponse(k50 = 3e12/g_rel_u, quanta = xx)
example_g_lows = PRresponse(k50 = 3e13/g_rel_u, quanta = xx)
example_u = PRresponse(k50 = 3e12, quanta = xx)

#set margins all around for individual plots
par(mar = c(4.5,4.5,4.5,4.5))
```

## Photoreceptor Response: Equal Absolute Sensitivity

First, we plot simulated response curves assuming the green and UV photoreceptors have equal absolute sensitivity. We also compute and overlay the log ratio of green to UV photoreceptor responses.

```{r plot_equal_sensitivity}
#open plot with green photoreceptor
plot(x = xx,
     y = example_g_equal,
     type = 'l',
     col = 'green',
     lwd = 5,
     log = 'x',
     ylim = c(0, 50),
     main = 'equal absolute sensitivity',
     xlab = 'UV intensity (photons/cm2/s)',
     ylab = 'photoreceptor response (mV)')
#add UV photoreceptor
lines(x = xx,
     y = example_u,
     col = 'magenta',
     lwd = 5)
#add reference lines
abline(h = c(0,0.5,1.0)*50,
       v = 10^c(11:15),
       col = 'gray')
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'))
#label curves
text(x = 10^c(12, 13.5, 13.7),
     y = c(25, 25, 30),
     labels = c('UV\nphotor.',
                'green\nphotor.',
                'log ratio'),
     col = c('magenta', 'green', 'purple4')
     )
#add log ratio
#open new plot
par(new = TRUE)
plot(x = xx,
     y = log(example_g_equal/example_u),
     type = 'l',
     ylim = c(-1.55, 0),
     log = 'x',
     lwd = 5,
     col = 'purple4',
     axes = FALSE,
     xlab = '',
     ylab = '')
#add axis
axis(side = 4,
     col = 'purple4')
#add axis label
mtext(side = 4,
      line = 2,
      text = 'log(green/UV)',
      col = 'purple4')
```

## Photoreceptor Response: Account for Sensitivity Difference

Next, we plot the response curves under a scenario where the green photoreceptor has a lower absolute sensitivity (higher semi-saturation constant $k_{50}$) than the UV photoreceptor.

```{r plot_lower_sensitivity}
plot(x = xx,
     y = example_g_lows,
     type = 'l',
     col = 'green',
     lwd = 5,
     log = 'x',
     ylim = c(0, 50),
     main = 'green lower absolute sensitivity',
     xlab = 'UV intensity (photons/cm2/s)',
     ylab = 'photoreceptor response (mV)')
lines(x = xx,
     y = example_u,
     col = 'magenta',
     lwd = 5)
#add reference lines
abline(h = c(0,0.5,1.0)*50,
       v = 10^c(11:15),
       col = 'gray')
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'))
#label curves
text(x = 10^c(12.2, 14.4, 13.7),
     y = c(25, 25, 25),
     labels = c('UV\nphotor.',
                'green\nphotor.',
                'log ratio'),
     col = c('magenta', 'green', 'purple4')
)
#add log ratio
#open new plot
par(new = TRUE)
plot(x = xx,
     y = log(example_g_lows/example_u),
     type = 'l',
     ylim = c(-4.0, 0),
     log = 'x',
     lwd = 5,
     col = 'purple4',
     axes = FALSE,
     xlab = '',
     ylab = '')

axis(side = 4,
     col = 'purple4')
mtext(side = 4,
      line = 2,
      text = 'log(green/UV)',
      col = 'purple4')
```

## Visualizing Figure 4

This combined two-panel figure presents:
1. The relative photoreceptor sensitivity profiles for both UV and green photoreceptors with integrated shaded areas for the UV and green LEDs.
2. The simulated response functions for both UV and green photoreceptors under the equal absolute sensitivity assumption, alongside the ratio of green to UV response on the right-hand axis.

```{r figure_4, fig.width=10, fig.height=5}
par(mfrow = c(1,2),
    mar = c(5,5,0,3))
plot(u_photopig,
     type = 'l',
     lwd = 5,
     col = 'magenta',
     xlim = c(300, 700),
     ylim = c(0,1),
     xlab = 'wavelength (nm)',
     ylab = 'photoreceptor sensitivity\n(relative)',
     )
lines(g_photopig,
      lwd = 5,
      col = 'green')

with(uPR_uLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('magenta',
                            alpha.f = 0.2),
          border = NA)
)
with(uPR_gLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('magenta',
                            alpha.f = 0.2),
          border = NA)
)

with(gPR_uLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('green',
                            alpha.f = 0.2),
          border = NA)
)
with(gPR_gLED,
  polygon(x = c(x, rev(x)),
          y = c(y, rep(0, length(y))),
          col = adjustcolor('green',
                            alpha.f = 0.2),
          border = NA)
)
abline(h = c(0,1.0),
       v = 100*c(3:7),
       col = 'gray')


plot(x = xx,
     y = example_g_equal,
     type = 'l',
     col = 'green',
     lwd = 5,
     log = 'x',
     ylim = c(0, 50),
     xlab = 'UV intensity (photons/cm2/s)',
     ylab = 'photoreceptor response (mV)')
lines(x = xx,
      y = example_u,
      col = 'magenta',
      lwd = 5)
#add reference lines
abline(h = c(0,0.5,1.0)*50,
       v = 10^c(11:15),
       col = 'gray')
#add LEDs
abline(v = c(1.2e12, 4e12, 4e14),
       col = c('magenta4', 'purple', 'magenta'))
#label curves
text(x = 10^c(12.2, 14.4, 13.7),
     y = c(25, 25, 25),
     labels = c('UV\nphotor.',
                'green\nphotor.',
                'ratio'),
     col = c('magenta', 'green', 'purple4')
)
#add ratio
#open new plot
par(new = TRUE)
plot(x = xx,
     y = example_g_equal/example_u,
     type = 'l',
     ylim = c(0, 1),
     log = 'x',
     lwd = 5,
     col = 'purple4',
     axes = FALSE,
     xlab = '',
     ylab = '')

axis(side = 4,
     col = 'purple4')
mtext(side = 4,
      line = 2,
      text = 'green/UV',
      col = 'purple4')
```
